---
title: "p8105_hw3_jl5549"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(p8105.datasets)
library(ggplot2)
```

Problem 1

```{r}
data("instacart")
```
The data contains `r {nrow(instacart)}` rows and `r {ncol(instacart)}` variables.  
Some key variables are shown following: order_dow: the day of the week on which the order was placed, order_hour_of_day: the hour of the day on which the order was placed, product_name: name of the product, aisle_id: aisle identifier, aisle: the name of the aisle.

```{r}
info_aisle= 
  instacart %>% 
  group_by(aisle) %>%
  summarize(n_items = n()) %>% 
  arrange(n_items)
n_aisle = nrow(info_aisle)
aisle_most_order = info_aisle[nrow(info_aisle),1]
```
There are `r {n_aisle}` aisles and the aisle the most items are orderd from is `r {aisle_most_order}`.
```{r}
#a plot showing the number of items ordered in each aisle
instacart %>% 
  group_by(aisle) %>% 
  summarize(n_items = n()) %>% 
  filter(n_items>10000) %>% 
  ggplot()+
  geom_col(aes(x = aisle, y = n_items))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1)) +
  labs(x = 'Aisle',
       y = 'Number of items ordered',
       title = 'The number of items ordered in each aisle')
```

It can be concluded that users order their products the most in aisles 'fresh friut' and 'fresh vegetables', and relatively less products in aisles 'oils vinegars' and 'butter'.
```{r}
#a table showing the three most popular items in three aisles
instacart %>%
  filter(aisle %in% c('baking ingredients', 'dog food care', 'packaged vegetables fruits')) %>%
   group_by(aisle,product_name) %>%  
  summarize(n = n()) %>% 
  filter(min_rank(desc(n)) < 4) 
 
```
Products ordered the most in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits” are Light Brown Sugar(499), Snack Sticks Chicken & Rice Recipe Dog Treats(30) and Organic Baby Spinach(9784) respectively. The number of products ordered in aisle 'packaged vegetables fruits' are generally more than other two aisles. 
```{r}
#a table showing the mean hour of the day
instacart %>%
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  filter( product_name %in% c('Pink Lady Apples', 'Coffee Ice Cream')) %>%
  pivot_wider(names_from = product_name,
              values_from = mean_hour) %>% 
  select(-order_dow)

```
Users normally order more coffe ice cream from Wednesday to Friday, and order the most Pink Lady Apples on Thursday. 

Problem2 

```{r}
data("brfss_smart2010")
mydata =brfss_smart2010 %>% 
    janitor::clean_names() %>% 
    filter(topic == "Overall Health",
           response %in% c("Poor", "Fair", "Good", "Very good", "Excellent")) %>%
   mutate(response = factor(response,levels = c("Poor", "Fair", "Good", "Very good", "Excellent"))) 
head(mydata)
```

```{r}
#states observed at 7 or more locations in 2002 and 2010 

mydata %>% 
  group_by(locationdesc, locationabbr, year) %>% 
  summarize(n=n()) %>%  
  filter(year %in% c(2002, 2010)) %>% 
  group_by(year, locationabbr) %>% 
  summarize(n = n()) %>% 
  filter(n > 6)
```

In 2002, states CT, FL, MA, NC, NJ, PA were observed at 7 or more locations.
In 2010, they are CA, CO, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX, WA.
```{r}
#a dataset that is limited to Excellent and a new variable created
excellent = mydata %>% 
  filter(response == 'Excellent') %>% 
  group_by(locationabbr, year) %>% 
  mutate(data_value_mean = mean(data_value))
head(excellent)
```
A “spaghetti” plot of this average value over time within a state is shown below:
```{r}
excellent %>% 
  ggplot()+
  geom_line(aes(x = year, y = data_value_mean,
                color = locationabbr, group = locationabbr))+
  scale_y_continuous(
    breaks = c(10, 20, 30), 
    limits = c(12, 30))
```

```{r}
#a two-panel plot showing distribution of data_value for responses among locations in NY State for the years 2006 and 2010
mydata %>%
  filter(locationabbr == 'NY',
         year %in% c(2006,2010)) %>% 
  ggplot()+
  geom_line(aes(x = response, y = data_value,
                color = locationdesc, 
                group = locationdesc ))+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  facet_grid(~year)
```

It can seen that 'Very good' is generally the most response in 2006 and 2010 across the counties in NY state.
Problem 3
```{r}
accel = read_csv(file = './accel_data.csv')

accel_tidy = accel %>%
  janitor::clean_names() %>% 
  mutate(weekday_vs_weekend = ifelse(day %in% c('Saturday', 'Sunday'), '0', '1')) %>% 
  select(weekday_vs_weekend, everything()) %>%  
  pivot_longer(5:ncol(.),
              names_to = 'activities',
              values_to = 'data',
              names_prefix = 'activity_')
head(accel_tidy)
```
There are `r {ncol(accel_tidy)}` variables existing, including following key variables:
weekday_vs_weekend: indicate whether the day monitored is a weedkay or a day of weekend. '1' indicates a weekday, otherwise a day in weekend
activities: nth minute of a 24-hour day starting at midnight
data: activity counts for yertain minute
There are `r {nrow(accel_tidy)}` observations.
```{r}
total = accel_tidy %>% 
  group_by(day_id, week, day) %>% 
  summarize(total = sum(data)) %>% 
  mutate(day = factor(day, levels =c('Monday', 'Tuesday', 
                                              'Wednesday', 'Thursday', 'Friday', 'Saturday', 'Sunday')))

```
According to the table created, the trend of activities for each day are not apparent.
```{r}
activity = 
  accel_tidy %>%
  group_by(week, day_id) %>% 
  mutate(day = factor(day, levels =c('Monday', 'Tuesday', 'Wednesday',
                                     'Thursday', 'Friday', 'Saturday', 'Sunday')),
         activities = as.integer(activities)) %>% 
  arrange(activities)
  
activity %>% 
  ggplot(aes(x = activities, y = data, color = day))+
  geom_line(alpha = 0.5)+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  scale_x_continuous(breaks  = seq(0, 1500, 60))
  
activity  
```
We can see from the graph that there are peaks at noon on Sunday and at night on Monday, Wednesday, Friday and Saturday.